### 2019-05-02 sections/semantics.tex
lambda -> \lambda
steps keep their right hand side unevaluated

In this section we formalise the semantics of the language constructs described in \cref{sec:language}.
We organise this by following the structure of the language.
Firstly, the task language is embedded in a simply typed $lambda$-calculus.
This requires a specification of the \emph{evaluation} of terms in the host language, and how it handles the task language.
Secondly, there are two ways to drive evaluation of task expressions, internally by the system itself, and externally by the user.
This is done in two additional semantics, one for the internal \emph{normalisation} of tasks, and another for the \emph{interaction} with the end user.

@@

\begin{figure}[h]
  \small
  \usemacro{G-Values-Compact}
  \caption{Value grammar} \label{fig:value-grammar}
\end{figure}

@@

The rules to evaluate expressions $e$ do not differ from standard work, except for the task constructs.
The evaluation rules for tasks can be deduced from the value grammar.
They are given in the appendix.
Most task constructors are strict in their arguments.
Only steps keep their left hand side unevaluated to delay side effects till the moment the step is taken.
The same holds for both branches of the external choice.

@@

% \begin{figure*}[b]
%   \small
%   \usemacro{O-Value-Compact} \hfill \usemacro{O-Failing-Compact} \hfill \usemacro{O-Inputs-Compact}
%   \caption{Values} \label{fig:observation-value}
% \end{figure*}

@@

\begin{figure}[h]
  \small
  \usemacro{O-Failing}
  \caption{Failing} \label{fig:observation-failing}
\end{figure}

@@

  \paragraph{Step}
  \begin{align*}
    \userule{S-ThenStay} \Quad
    & \userule{S-ThenFail} \Break
    & \userule{S-ThenCont}
  \end{align*}

  \paragraph{Choose}
  \begin{align*}
    \userule{S-OrLeft} \Quad
    & \userule{S-OrRight} \Break
    & \userule{S-OrNone}
  \end{align*}

  \paragraph{Ready}
  \begin{gather*}
    \userule{S-Edit} \Quad \userule{S-Fill} \Quad \userule{S-Update} \Quad
    \userule{S-Fail} \Quad \userule{S-Xor}
  \end{gather*}

@@

\begin{figure}[h]
  \small
  \usemacro{G-Inputs-Compact}
  \caption{Input grammar} \label{fig:input-grammar}
\end{figure}

\begin{figure}[h]
  \small
  \usemacro{O-Inputs}
  \caption{Inputs} \label{fig:observation-input}
\end{figure}

@@

  \paragraph{Passing}
  \begin{gather*}
    \userule{H-PassThen} \Quad \userule{H-FirstAnd} \Quad \userule{H-FirstOr} \Break
    \userule{H-PassNext} \Quad \userule{H-SecondAnd} \Quad \userule{H-SecondOr}
  %  \userule{H-Appoint}
  \end{gather*}

>>>>

In this section we formalise the semantics of the language constructs described in \cref{sec:language}.
We organise this by following the structure of the language.
Firstly, the task language is embedded in a simply typed $\lambda$-calculus.
This requires a specification of the \emph{evaluation} of terms in the host language, and how it handles the task language.
Secondly, there are two ways to drive evaluation of task expressions, internally by the system itself, and externally by the user.
This is done in two additional semantics, one for the internal \emph{normalisation} of tasks, and another for the \emph{interaction} with the end user.

@@

\begin{figure}[h]
  \small
  \usemacro{G-Values-Minimal}
  \caption{Value grammar} \label{fig:value-grammar}
\end{figure}

@@

The rules to evaluate expressions $e$ do not differ from standard work, except for the task constructs.
The evaluation rules for tasks can be deduced from the value grammar.
They are given in the appendix.
Most task constructors are strict in their arguments.
Only steps keep their right hand side unevaluated to delay side effects till the moment the step is taken.
The same holds for both branches of the external choice.

@@

% \begin{figure*}[b]
%   \small
%   \usemacro{O-Value-Minimal} \hfill \usemacro{O-Failing-Minimal} \hfill \usemacro{O-Inputs-Minimal}
%   \caption{Values} \label{fig:observation-value}
% \end{figure*}

@@

\begin{figure}[h]
  \small
  \usemacro{O-Failing-Compact}
  \caption{Failing} \label{fig:observation-failing}
\end{figure}

@@

  \paragraph{Step}
  \begin{gather*}
    \userule{S-ThenStay} \Break
    \userule{S-ThenFail} \Break
    \userule{S-ThenCont}
  \end{gather*}

  \paragraph{Choose}
  \begin{gather*}
    \userule{S-OrLeft} \Break
    \userule{S-OrRight} \Break
    \userule{S-OrNone}
  \end{gather*}

  \paragraph{Ready}
  \begin{gather*}
    \userule{S-Edit} \Quad \userule{S-Fill} \Quad \userule{S-Update} \Break
    \userule{S-Fail} \Quad \userule{S-Xor}
  \end{gather*}

@@

\begin{figure}[h]
  \small
  \usemacro{G-Inputs-Minimal}
  \caption{Input grammar} \label{fig:input-grammar}
\end{figure}

\begin{figure}[h]
  \small
  \usemacro{O-Inputs-Compact}
  \caption{Inputs} \label{fig:observation-input}
\end{figure}

@@

  \paragraph{Passing}
  \begin{gather*}
    \userule{H-PassThen} \Quad \userule{H-PassNext} \Break
    \userule{H-FirstAnd} \Quad \userule{H-SecondAnd} \Break
    \userule{H-FirstOr}  \Quad \userule{H-SecondOr}
  %  \userule{H-Appoint}
  \end{gather*}
